MLH1 Data Report

### 2019-12-19 doc/Main_Report_Results.rmd
updates and notes for the mixed model results

setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
#setwd("C:/Users/April/Desktop/MLH1repo")
#load(file="data/MLH1/MLH1_data_setup.RData")

@@

load(file="data/MLH1/MLH1_data_setup_12.3.19.RData")#removed exclude mice
#load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")#added batch17
#load(file="data/MLH1/MLH1_data_setup_11.11.19.RData")#deleted DUP image
#load(file="data/MLH1/MLH1_data_setup_9.26.19.RData")#added batch16
#load(file="data/MLH1/MLH1_data_setup_8.29.19.RData") #
#load(file="data/MetaData.RData") #this was replacing most updated data file

@@

#AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[!( AP_mouse_table_w.Ages$strain.x =="F1"), ]

@@

# Distributions of MLH1 counts across Samples

The histograms below are for total / pooled cells (for now, maybe compare with HQ data to see what changes)

```{r MLH1.histogram.2421, echo=FALSE, fig.caption= "histograms", warning=FALSE, fig.align='center'}

#gg <- data.frame(weight=rep(x,nrow(df)),df)
#gg$y <- with(gg,dnorm(x,mean,sd))
#gg$y <- gg$y * aggregate(weight~sex, weight,length)$weight * #diff(range(weight$weight))/30

mean.MLH1.obs <- ddply(.data=MLH1_data, 
                 .(category),
                 summarize, 
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)

histM <- ggplot(data=MLH1_data,(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ category)+ggtitle("Pooled Cell Histograms")

#when I try to add the mean lines, error returned that it can't find the adj_MLH1 ....?
#histM <- histM +  geom_line(data=mean.MLH1.obs, aes(y=mean.MLH1.obs$mean.MLH1))

#ad
#  geom_smooth(method = lm, se = FALSE
 # geom_smooth(method = lm, formula = mean.MLH1.obs$mean.MLH1 ~ mean.MLH1.obs$category, se = FALSE)

histM

#playing with expected posisson plot...
#function = glm(y ~ x, family = poisson)
#add Poission curve behind
#geom_smooth(method = lm, formula = y ~ splines::bs(x, 3), se = FALSE)
#not sure how to add Expected poisson curve

Most histograms of pooled MLH1 counts are all close to normal. Except PWD male, LEW female, SKIVE male, SPIC male.
check if there is a quality effect.

```{r MLH1.histogram.2, echo=FALSE}
#remake the pooled MLH1 histgoram plot from above but facet by quality

#PWD male, MSM male
#LEW female, WSB female
mean.MLH1.obs_qual <- ddply(.data=MLH1_data, 
                 .(category, quality),
                 summarize, 
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)
#what's the point of the above table? -- add vlines?

#I think I should do some t-tests across the groups of quality bins to better asess the amount of quality effect

histM_qual.KAZ_male <- ggplot(data=MLH1_data[MLH1_data$category == "KAZ male",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("KAZ male Pooled Cell Histograms by quality")+ geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "KAZ male"],color="red")
histM_qual.KAZ_male

histM_qual.PWD_male <- ggplot(data=MLH1_data[MLH1_data$category == "PWD male",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("PWD male Pooled Cell Histograms by quality")+ geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "PWD male"],color="red")
histM_qual.PWD_male

histM_qual.MSM_male <- ggplot(data=MLH1_data[MLH1_data$category == "MSM male",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("MSM male Pooled Cell Histograms by quality")+geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "MSM male"], color="red")
histM_qual.MSM_male

histM_qual.WSB_female <- ggplot(data=MLH1_data[MLH1_data$category == "WSB female",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("WSB female Pooled Cell Histograms by quality")+ geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "WSB female"],color="red")
histM_qual.WSB_female

histM_qual.LEW_female <- ggplot(data=MLH1_data[MLH1_data$category == "LEW female",],(aes(x=adj_nMLH1.foci))) + xlim(c(16, 38)) + geom_histogram(stat = "bin", binwidth = 1) + facet_wrap(~ quality, ncol = 5)+ggtitle("LEW female Pooled Cell Histograms by quality")+ geom_vline(xintercept=mean.MLH1.obs_qual$mean.MLH1[mean.MLH1.obs_qual$category == "LEW female"],color="red")
histM_qual.LEW_female

#plot.pvals.L <- ggplot(bunchOpvals.L, aes(bunchOpvals.L) )+geom_histogram(bins = 50)+ ggtitle("1000 rep t-tests sample size 25, High vs Low Leptotene")+geom_vline(xintercept=ttest.HighLow.L$p.value, linetype="dotted", color="red")

#when I try to add the mean lines, error returned that it can't find the adj_MLH1 ....?
#histM <- histM +  geom_line(data=mean.MLH1.obs, aes(y=mean.MLH1.obs$mean.MLH1))

#ad
#  geom_smooth(method = lm, se = FALSE
 # geom_smooth(method = lm, formula = mean.MLH1.obs$mean.MLH1 ~ mean.MLH1.obs$category, se = FALSE)

@@

@@

$$mouse .mean CO ~=~ subsp * sex + rand(strain* sex)$$

$$mouse .var CO ~=~ subsp * sex + rand(strain* sex)$$

@@

```{r mixed.models.setup, echo=FALSE, warning=FALSE}

#why merged 
AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x

#main file, make everything unordered factors
AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]

@@

#make dummy MOLF data, average of Musc females?
#no MOLF male data?
female.musc.cells <- MLH1_data[( MLH1_data$sex == "female") & (MLH1_data$subsp == "Musc" ), ]
table(female.musc.cells$quality)

@@

#female.musc.mice.table$mouse <- "1jan20_MOLF_f99"
#female.musc.mice.table$DOB <- as.Date("1/1/2020")

#female.musc.mice.table$strain <- "MOLF"
#female.musc.mice.table$subsp <- "Musc"
#female.musc.mice.table$species <- "M.musculus"

#female.musc.mice.table$female.stage <- "emb"
#female.musc.mice.table$age.days <-""
#female.musc.mice.table$age.weeks <- ""
#female.musc.mice.table$ID <- ""
#female.musc.mice.table$notes <- ""
#female.musc.mice.table$euth.date <- ""

#DF.HetC.MixedModel <- rbind(DF.HetC.MixedModel, female.musc.mice.table)

DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
#remove musc strains
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
#remove PERC, MOLF, CZECH, (no females)
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "MOLF",]

DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)
DF.HetC.MixedModel <- DF.HetC.MixedModel[!(is.na(DF.HetC.MixedModel$mouse) | DF.HetC.MixedModel$mouse==""), ]

#main df with variables coded as chr
DF.HetC.MixedModel <- DF.HetC.MixedModel
DF.HetC.MixedModel$subsp <- factor(DF.HetC.MixedModel$subsp, ordered = FALSE)
DF.HetC.MixedModel$sex <- factor(DF.HetC.MixedModel$sex, ordered = FALSE)
DF.HetC.MixedModel$strain <- factor(DF.HetC.MixedModel$strain, ordered = FALSE)

#mean_co.HQ  (this might remove some of the MOLF males)
DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[DF.HetC.MixedModel$Ncells > 9,]
DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$age.weeks < 25,]
DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse) | DF.HetC.MixedModel.HQ$mouse==""), ]

#mean_co.HQ.mol
DF.HetC.MixedModel.HQ.mol <- DF.HetC.MixedModel.HQ

@@

For mouse averages of MLH1 counts, only strain with observations for both sexes and are from Mus species. (maybe I'll look at Species level with spicilegus.?)

subspecies levels: 2, Musc and Domesticus

strain levels: Dom: 3, Musc: 4

@@

These are the List of mixed models I built

1. mean_co (chr coded, instead of factors)

2. mean_co.HQ

3. mean_co with Mol (Mol coded as a distinct subsp)

4. sub-sampling (not sure if I made this one)

```{r HQ.MixedModel.1, echo=FALSE}
#table(DF.HetC.MixedModel$strain, DF.HetC.MixedModel$sex)

#mean_co.HQ  (this might remove some of the MOLF males)
DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[DF.HetC.MixedModel$Ncells > 9,]
DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$age.weeks < 25,]
DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse) | DF.HetC.MixedModel.HQ$mouse==""), ]

#mean.co.HQ
HetC.MixedModel.HQ <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )

#summary(HetC.MixedModel.HQ)
#ranef(HetC.MixedModel.HQ)
```

```{r MixedModel.2, echo=FALSE}
#reduced models
Reduced.strain.HQ <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random = ~1|strain)
#summary(Reduced.strain.HQ)
#ranef(Reduced.strain.HQ)
```

```{r MixedModel.2.2, echo=FALSE}

Reduced.all.HQ <- gls(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ)
#summary(Reduced.all.HQ)
#ranef(Reduced.all.HQ)
```

```{r MixedModel.3, echo=FALSE}
#strain = fixed
strain.fixed.full <- lm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)
#summary(strain.fixed.full)
#when strain added as fixed, MSM, PWD and WSB are signifcant, subsp*sex is most significant

@@

@@

strain.fixed <- lm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ)
#summary(strain.fixed)

#maybe this WSB effec is just in the HQ data set?
#anova for comparing reduced and full models -- 1 of the models should be strain as fixed instead
full2strain <- anova(HetC.MixedModel.HQ, Reduced.strain.HQ)#slightly significant

@@

full2noRandom <- anova(HetC.MixedModel.HQ, Reduced.all.HQ)# < .0001

strain2noRandom <- anova(Reduced.strain.HQ, Reduced.all.HQ)#NS
```

```{r try.blah, echo=FALSE}
#library(jtools)
#states <- as.data.frame(state.x77)
#fit <- lm(Income ~ Frost + Illiteracy + Murder, data = states)
#summ(fit)
#this is a nice display package for lm

res.table <- as.data.frame(coef(summary(Reduced.all.HQ)))
res.table
```

#plot of variance
var.boxplot <- ggplot(DF.HetC.MixedModel.HQ, aes(x = as.factor(mouse), y = var, fill= strain,alpha=0.5))+ geom_point()+
 ggtitle("Boxplots of MLH1 distributions") + theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
 ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +   scale_fill_manual(values=colors_of_strains)
var.boxplot <- var.boxplot   + facet_wrap(sex ~ strain, scales="free")

cV.male <- ggplot(DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$sex == "male",], aes(x = as.factor(mouse), y = cV, fill= strain,alpha=0.5))+ geom_point()+
 ggtitle("Male variance levels") + theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  scale_fill_manual(values=colors_of_strains)+  ylim( c( 5, 20) )

```{r BaseMixedModel, echo=FALSE, eval=FALSE, include=FALSE}
#https://rcompanion.org/handbook/G_03.html
#testing random effects within lme
#strain as fixed
newnew <- lme(mean_co ~ sex * strain, data=DF.HetC.MixedModel )

HetC.MixedModel.mean_co <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel, random=list(strain=pdDiag(~sex) ) )

R.strain.HetC.MixedModel.mean_co <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel, random = ~1|strain)#musc-male fixed effect is now significant, and male

#CONTRAST models

#base vs strain
fullVstrain <- anova(HetC.MixedModel.mean_co, R.strain.HetC.MixedModel.mean_co)# <.0001

strain.vs.non <- anova(R.strain.HetC.MixedModel.mean_co, full.Reduced.HetC.MixedModel.mean_co)#<.0001

#fullreduce vs strain

#can't distinguish which random effects are more significant -- both tests are significant

#random effects for base model
ranef(HetC.MixedModel.mean_co)
#random effect for just strain
ranef(R.strain.HetC.MixedModel.mean_co)

#reduced model's PWD radom effects is much smaller
#the full model's largest random effects are for PWD and MSM (and MOLF, KAZ for negative)

#fixed effects from full
anova(HetC.MixedModel.mean_co)#NS

#Check how the fixed effects change across these

HetC.MixedModel.mean_co2 <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel)
#lmer(y~x1 + x2 + (1|g1), data=pbDat)

lmer.MM <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel, REML = TRUE)

@@

HetC.MM.strain.fixed2 <- lm(mean_co ~ (subsp/strain)*sex, data= DF.HetC.MixedModel )
summary(HetC.MM.strain.fixed2)
#(WSB and G have the most observations, maybe this is biasing the results (for WSB?))

@@

```{r MolossinusMixedModel, echo=FALSE, eval=FALSE, include=FALSE}
#base model with molossinus level ... why am I making molossinus model?

DF.HetC.MixedModel.mol <- DF.HetC.MixedModel
str(DF.HetC.MixedModel.mol)#subsp is factor with 2 levels

DF.HetC.MixedModel.mol$subsp <- as.character(DF.HetC.MixedModel.mol$subsp)

#turn to chr first?
#apparently ifelse doesn't retain class
library(data.table)

DF.HetC.MixedModel.mol$subsp <- if_else(grepl("MSM", DF.HetC.MixedModel.mol$strain), "Mol", 
        if_else(grepl("MOLF", DF.HetC.MixedModel.mol$strain), "Mol", DF.HetC.MixedModel.mol$subsp))

str(DF.HetC.MixedModel.mol)
DF.HetC.MixedModel.mol$subsp <- factor(DF.HetC.MixedModel.mol$subsp, ordered = FALSE)

#these remove ALL male MSM!
table(DF.HetC.MixedModel.HQ$strain, DF.HetC.MixedModel.HQ$sex)

#mean.co
HetC.MixedModel.mol <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.mol, random=list(strain=pdDiag(~sex) ) )

summary(HetC.MixedModel.mol)
ranef(HetC.MixedModel.mol)

#mreducing the model (reducing the random effects)
Reduced.strain.mol <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.mol, random = ~1|strain)
summary(Reduced.strain.mol)
ranef(Reduced.strain.mol)

Reduced.all.mol <- gls(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.mol)
summary(Reduced.all.mol)

#base vs mol model (this only makes sense if these are nested models)
anova(HetC.MixedModel.mean_co, HetC.MixedModel.mol)

@@

Issue presented with is that we can see evolution across the strains, but it isn't predicted from mixed models by strain.
make artificial data / models, like human labels

## Mixed Model for variance

```{r HQ.var, echo=FALSE}
#modeling within mouse variance

@@

summary(HetC.MixedModel.HQ.var_co)
ranef(HetC.MixedModel.HQ.var_co)

HQjust.strain.var_co <- lme(cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random = ~1|strain)

summary(HQjust.strain.var_co)

HQfull.Reduced.cv_co <- gls(cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ)
summary(HQfull.Reduced.cv_co)

@@

strain2noRandom.CV <- anova(HQjust.strain.var_co, HQfull.Reduced.cv_co)# 0.0469

>>>>

#setwd("~./MLH1repo/doc/")
setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
#setwd("C:/Users/April/Desktop/MLH1repo")
#load(file="data/MLH1/MLH1_data_setup.RData")

@@

load(file="~./data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
#load(file="data/MLH1/MLH1_data_setup_12.3.19.RData")#removed exclude mice
#load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")#added batch17
#load(file="data/MLH1/MLH1_data_setup_11.11.19.RData")#deleted DUP image
#load(file="data/MLH1/MLH1_data_setup_9.26.19.RData")#added batch16
#load(file="data/MLH1/MLH1_data_setup_8.29.19.RData") #
#load(file="data/MetaData.RData") #this was replacing most updated data file

@@

AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[!( AP_mouse_table_w.Ages$strain.x =="F1"), ]

@@

@@

*cast female data came from another publication

@@

$$mouse .mean CO ~=~ subsp * sex + rand(strain* sex)  + \varepsilon$$

$$mouse .var CO ~=~ subsp * sex + rand(strain* sex) + \varepsilon $$

@@

```{r mixed.models.setup, echo=FALSE, warning=FALSE}
#https://rcompanion.org/handbook/G_03.html
#testing random effects within lme

#notes for testing random effects

#variance is still chr
AP_mouse_table_w.Ages$var <- as.numeric(AP_mouse_table_w.Ages$var)

#use the merged DF to filter by age 
AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x
#main file, make everything unordered factors
AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)

DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]

@@

#why did I male just a female table?
female.musc.cells <- MLH1_data[( MLH1_data$sex == "female") & (MLH1_data$subsp == "Musc" ), ]

@@

#REMOVE EXTRA STRAINS
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
#remove musc strains
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
#remove PERC, MOLF, CZECH, (no females)
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]

DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)#remove levels not used

#remove blank rows (which are these?)
#DF.HetC.MixedModel <- DF.HetC.MixedModel[!(is.na(DF.HetC.MixedModel$mouse) | #DF.HetC.MixedModel$mouse==""), ]
#151 mice, no rows with nas for values

#main df with variables coded as chr
DF.HetC.MixedModel <- DF.HetC.MixedModel
DF.HetC.MixedModel$subsp <- factor(DF.HetC.MixedModel$subsp, ordered = FALSE)
DF.HetC.MixedModel$sex <- factor(DF.HetC.MixedModel$sex, ordered = FALSE)
DF.HetC.MixedModel$strain <- factor(DF.HetC.MixedModel$strain, ordered = FALSE)
```

```{r HQ.dataset, echo=FALSE}

#also make a HQ dataset, tighter window for ages,

#mean_co.HQ  (this might remove some of the MOLF males)
DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[DF.HetC.MixedModel$Ncells > 9,]
DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$age.weeks < 25,]
DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse) | DF.HetC.MixedModel.HQ$mouse==""), ]
```

```{r Mol.dataset, echo=FALSE}
#mean_co.HQ.mol
DF.HetC.MixedModel.HQ.mol <- DF.HetC.MixedModel.HQ

@@

Seting up the data set for using the mixed model (base data set, HQ dataset)

- only Mus musculus strains with sex matched observations

- Unorder factors

- male ages within range

Currently I ran through 3 datasets: 1) base, 2) HQ and 3) Molf (from HQ)

@@

For mouse averages of MLH1 counts, only strain with observations for both sexes and are from Mus species.

subspecies levels: 2(3), Musc and Domesticus (Molossinus)

strain levels: Dom: 3, Musc: 4  (Mol: 2)

## Mean CO Count

$$mouse .mean CO ~=~ subsp * sex  + rand(strain*sex) + \varepsilon$$

```{r HQ.MixedModel.1, echo=FALSE, digits = 2}
#table(DF.HetC.MixedModel$strain, DF.HetC.MixedModel$sex)
#mean_co.HQ  (this might remove some of the MOLF males)

HetC.MixedModel.HQ <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )

as.data.frame( round(  summary(HetC.MixedModel.HQ)$tTable, 6 ) )

#ranef(HetC.MixedModel.HQ)
#which is the test for random effects

Above is the table of coefficients for MM.1, HQ dataset, lme(), random effects strain and strain specific sex effect.

$$mouse .mean CO ~=~ subsp * sex  + rand(strain*sex) + \varepsilon$$

```{r MM.1.Mol, echo=FALSE}

HetC.MixedModel.HQ.mol <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ.mol, random=list(strain=pdDiag(~sex) ) )

as.data.frame( round(   summary(HetC.MixedModel.HQ.mol)$tTable, 6) )

Above is the table of coefficients for MM.1.mol, molssinus coded as seperate subspecies, lme(), random effects strain and strain specific sex effect.

$$mouse .mean CO ~=~ subsp * sex  + rand(1|strain) + \varepsilon$$

```{r MixedModel.2, echo=FALSE}
#reduced models
Reduced.strain.HQ <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random = ~1|strain)
#summary(Reduced.strain.HQ)

as.data.frame(round(summary(Reduced.strain.HQ)$tTable, 6))
#add a aption to this code chunk
#ranef(Reduced.strain.HQ)

Above is the table of coefficients for MM.2, HQ dataset, lme(), random effects coded without sex specific effect.
This is model has sex and interaction coefficient is significant.

> MM.2.Mol

```{r MixedModel.2.mol, echo=FALSE}
#reduced models
Reduced.strain.HQ.mol <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ.mol, random = ~1|strain)
#summary(Reduced.strain.HQ)

as.data.frame(round(summary(Reduced.strain.HQ.mol)$tTable, 6))
#add a aption to this code chunk
#ranef(Reduced.strain.HQ.mol)

The model for MM.2.Mol has similar results as the above model, with just 2 subsps
This model has significant coefficients for sex and the 2 interaction effects (sex:musc, sex:mol)

> MM2.2

$$mouse .mean CO ~=~ subsp * sex  + \varepsilon$$

```{r MixedModel.2.2, echo=FALSE}
Reduced.all.HQ <- gls(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ)

as.data.frame( round(summary(Reduced.all.HQ)$tTable, 6))
#summary(Reduced.all.HQ)
#ranef(Reduced.all.HQ)
```

Above is the table of coefficients for MM.2.2, HQ dataset, gls(), No random effects.  
I think this gls has sex and interaction term as significant.

> MM3

$$mouse .mean CO ~=~ subsp * sex * strain + \varepsilon$$

```{r MixedModel.3, echo=FALSE}
#strain = fixed
strain.fixed.full <- lm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)
#library(xtable)

#this table of results is broken -- might be too long -- lm
#lm has different names
as.data.frame(round( summary(strain.fixed.full)$coefficients, 6))

#summary(strain.fixed.full)
#when strain added as fixed, MSM, PWD and WSB are signifcant, subsp*sex is most significant

@@

Above is the table of coefficients for MM.3, HQ dataset, lm function, sex, strain and subsp are interacting fixed effects.

Strain G is significant (p=0.0005),

interaction terms of sex and strain for (G, MSM, PWD).

How are these models different from the last one -- MM4 is missing subsp, so only sex * strain effects.

> MM.4

$$mouse .mean CO ~=~ sex * strain + \varepsilon$$

@@

strain.fixed <- lm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ)

#below table is broken
as.data.frame( round( summary(strain.fixed)$coefficients, 6))

#summary(strain.fixed)
#maybe this WSB effec is just in the HQ data set?
```

Above is the table of coefficients for **MM.4**, HQ dataset, lm function, just sex and strain, no subspecies.

Significant factors are 
strainG (p=0.0005), strainMSM and strainPWD (very slight),
sex:G (p=0.001),
sex:MSM, Sex:PWD (very slight p=.05, p=0.3),

#### ANOVA, mean CO count full vs reduced

Use Anova to compare full vs reduced models
the reduced models will be those missing either, strain, sex, or subsp.

```{r anova.full.v.reduced, echo=FALSE, eval=FALSE}

#strain as fixed
#CONTRAST models
#newnew <- lme(mean_co ~ sex * strain, data=DF.HetC.MixedModel )

HetC.MixedModel.mean_co <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel, random=list(strain=pdDiag(~sex) ) )
R.strain.HetC.MixedModel.mean_co <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel, random = ~1|strain)#musc-male fixed effect is now significant, and male

#base vs strain
fullVstrain <- anova(HetC.MixedModel.mean_co, R.strain.HetC.MixedModel.mean_co) #<.0001

#show scientific vs normal
#vv <- fullVstrain$`p-value`
#round(vv[2], 8)
#format(vv[2], scientific=F)

strain.vs.non <- anova(R.strain.HetC.MixedModel.mean_co, full.Reduced.HetC.MixedModel.mean_co) #<.0001

#fullreduce vs strain
#can't distinguish which random effects are more significant -- both tests are significant

#anova for comparing reduced and full models -- 1 of the models should be strain as fixed instead
full2strain <- anova(HetC.MixedModel.HQ, Reduced.strain.HQ)#slightly significant

@@

full2noRandom <- anova(HetC.MixedModel.HQ, Reduced.all.HQ) # < .0001

strain2noRandom <- anova(Reduced.strain.HQ, Reduced.all.HQ) #NS
```

```{r MolossinusMixedModel, echo=FALSE, eval=FALSE, include=FALSE}
#base model with molossinus level ... why am I making molossinus model?
#mean.co
HetC.MixedModel.mol <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.mol, random=list(strain=pdDiag(~sex) ) )

#mreducing the model (reducing the random effects)
Reduced.strain.mol <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.mol, random = ~1|strain)

Reduced.all.mol <- gls(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.mol)

#base vs mol model (this only makes sense if these are nested models)
anova(HetC.MixedModel.mean_co, HetC.MixedModel.mol)
```

@@

HetC.MM.strain.fixed2 <- lm(mean_co ~ (subsp/strain)*sex, data= DF.HetC.MixedModel )

#summary(HetC.MM.strain.fixed2)
#(WSB and G have the most observations, maybe this is biasing the results (for WSB?))

@@

#### Summary of mean COs

MM.2 has significant coefficients, sex and sex*subsp interaction.

MM.2.Mol, the sex and interaction term (subsp*sex) are significant.

MM.4 the strains which are significant are G, PWD and MSM. These 3 strains also have significant interaction (sex*strain) terms.

# Variance in CO count

The **biological significance** of within mouse variance in CO per cell is...

I can show the variance boxplots?  Not sure if it's very helpful.
Use the same Datasets HQ and mol for the mean CO counts. Use both var and cV as dependant variables

DF.HetC.MixedModel.HQ.mol
DF.HetC.MixedModel.HQ

#plot of variance
var.boxplot <- ggplot(DF.HetC.MixedModel.HQ, aes(x = as.factor(mouse), y = var, fill= strain,alpha=0.5))+ geom_point()+
 ggtitle("Boxplots of MLH1 distributions") + theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
 ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +   scale_fill_manual(values=colors_of_strains)
var.boxplot <- var.boxplot   + facet_wrap(sex ~ strain, scales="free")
#this plot shows variance as points/single values

cV.male <- ggplot(DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$sex == "male",], aes(x = as.factor(mouse), y = cV, fill= strain,alpha=0.5))+ geom_point()+
 ggtitle("Male cV levels") + theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  scale_fill_manual(values=colors_of_strains)+  ylim( c( 5, 20) )

@@

$$mouse .cV CO ~=~ subsp * sex + rand(strain*sex)+ \varepsilon$$

```{r cV.model.setup, echo=FALSE}
#modeling within mouse variance

@@

#summary(HetC.MixedModel.HQ.var_co)

as.data.frame( round(summary(HetC.MixedModel.HQ.var_co)$tTable, 6))

#ranef(HetC.MixedModel.HQ.var_co)
```

lme function, Cv modeled. Only sex is significant.

```{r cv.mod1.mol, echo=FALSE}

HetC.MixedModel.HQ_mol.cV_co <- lme(cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ.mol, random=list(strain=pdDiag(~sex) ) )

as.data.frame( round(summary(HetC.MixedModel.HQ_mol.cV_co)$tTable, 6))

$$mouse .cV CO ~=~ subsp * sex + rand(1|strain)+ \varepsilon$$

```{r cV.models, echo=FALSE}
HQjust.strain.var_co <- lme(cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random = ~1|strain)

as.data.frame( round(summary(Reduced.all.HQ)$tTable, 6))

#summary(HQjust.strain.var_co)
```

Sex and sex*subsp interaction effect are significant.

```{r cV.mod.mol, echo=FALSE}

HQ.mol.cV_co <- lme(cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ.mol, random = ~1|strain)

as.data.frame( round(summary(HQ.mol.cV_co)$tTable, 6))

Sex is significant, when mol is a seperate subspecies interaction is no longer significant.

$$mouse .cV CO ~=~ subsp * sex + \varepsilon$$
```{r more.cV.CO, echo=FALSE}
HQfull.Reduced.cv_co <- gls(cV ~ subsp * sex, data=DF.HetC.MixedModel.HQ)
as.data.frame( round(summary(HQfull.Reduced.cv_co)$tTable, 6))
#summary(HQfull.Reduced.cv_co)

@@

strain2noRandom.CV <- anova(HQjust.strain.var_co, HQfull.Reduced.cv_co)# 0.0469
```

$$mouse .var CO ~=~ subsp * sex + rand(strain*sex) + \varepsilon$$

#there's 1 mouse with just 1 cell remove
DF.HetC.MixedModel <- DF.HetC.MixedModel[!(is.na(DF.HetC.MixedModel$var) | DF.HetC.MixedModel$var==""), ]
DF.HetC.MixedModel$var <-  as.numeric(DF.HetC.MixedModel$var)
